---
title: "STA303 - Final Project"
author: "Cristina Su Lam"
date: "2024-03-09"
output: pdf_document
---
Load packages
```{r}
suppressMessages({
  suppressWarnings({
    library(knitr)
    library(gmodels)
    library(readr)
    library(magrittr) 
    library(dplyr)
    library(tidyr)
    library(glmnet)
    library(dtplyr)
    library(glmnet)
    library(MASS)
    library(rms)
    library(pROC)
  })
})
```

# 1. Read the Dataset & Drop Missing Values
```{r, warning=FALSE}
customer_booking <- read_csv("customer_booking.csv")
 
# Rename the category "CircleTrip" to "RoundTrip" in the "trip_type" variable
customer_booking <- customer_booking %>%
  mutate(trip_type = ifelse(trip_type == "CircleTrip", "RoundTrip", trip_type))

# Filter the dataset to only include customers from Australia
customer_booking_malaysia <- customer_booking[customer_booking$booking_origin == "Malaysia", ]

# Remove the column 'booking_origin'
customer_booking_malaysia <- customer_booking_malaysia[, -which(names(customer_booking_malaysia) == "booking_origin")]

# Check the first few rows of the dataset
head(customer_booking_malaysia)
```

```{r}
# Check for missing values in the entire data set
missing_values <- any(is.na(customer_booking_malaysia))

# Print the result
if (missing_values) {
  print("The dataset contains missing values.")
} else {
  print("The dataset does not contain missing values.")
}
```

# 2. Logistic Regression Model (All variables)
```{r, warning=FALSE}
# Fit logistic regression model using original data set
logit_model <- glm(booking_complete ~ sales_channel + trip_type + purchase_lead + length_of_stay + flight_hour +
                      wants_preferred_seat + num_passengers + flight_day + route + wants_extra_baggage +
                      wants_in_flight_meals + flight_duration, 
                    family = binomial(link = logit), 
                    data = customer_booking_malaysia)
```

# 3. Model Selection 

Perform stepwise AIC selection
```{r}
sel.var.aic <- step(logit_model, trace = 0, k = 2, direction = "both") 
select_var_aic <- attr(terms(sel.var.aic), "term.labels")   
select_var_aic
```

Perform stepwise BIC selection
```{r}
sel.var.bic <- step(logit_model, trace = 0, k = log(nrow(customer_booking_malaysia)), direction = "both") 
select_var_bic <- attr(terms(sel.var.bic), "term.labels")   
select_var_bic
```

LASSO Method 
```{r}
set.seed(1007928566)

# x contains the predictors and y contains the response variable
x <- model.matrix(booking_complete ~ ., data = customer_booking_malaysia)[,-1]
y <- customer_booking_malaysia$booking_complete

# Fit the model 
fit <- glmnet(x, y, family = "binomial")

# Make predictions for all observations
predictions <- predict(fit, newx = x, type = "class", s = c(0.05, 0.01))

# Evaluate model performance
cv.out <- cv.glmnet(x, y, family = "binomial", type.measure = "class", alpha = 1)

# Plot the cross-validation results
plot(cv.out)

# Get the best lambda value
best.lambda <- cv.out$lambda.1se

# Get the coefficients at the selected lambda
co <- coef(cv.out, s = "lambda.1se")

# Threshold for variable selection
thresh <- 0.00

# Select variables
inds <- which(abs(co) > thresh)
variables <- row.names(co)[inds]
sel.var.lasso <- variables[!(variables %in% '(Intercept)')]
sel.var.lasso
```

# 6. Model Diagnostics & Validation

## 6.1.1 Stepwise AIC Model

Fit Logistic Regression Model with AIC selection
```{r, warning=FALSE}
aic.logit <- glm(booking_complete ~ sales_channel + length_of_stay + flight_duration + num_passengers + 
                   wants_extra_baggage + wants_preferred_seat + wants_in_flight_meals, family = binomial(link = logit),
                 data = customer_booking_malaysia)
summary(aic.logit)
```

Checking influential points
```{r}
# DFBETAS
df.aic <- dfbetas(aic.logit)
n <- nrow(customer_booking_malaysia)
beta_cut <- 2 / sqrt(n)
influential_points <- apply(abs(df.aic) > beta_cut, 1, any)
sum(influential_points)
```

Checking outliers
```{r}
ri.aic <- rstandard(aic.logit)
outliers_obs <- which(abs(ri.aic) > 2)
length(outliers_obs)
```

VIF to check for multicollinearity
```{r}
vif(aic.logit)
```

Cross-Validation and Calibration
```{r}
set.seed(1007928566)
lrm.aic <- lrm(booking_complete ~ ., 
                 data = customer_booking_malaysia[, which(colnames(customer_booking_malaysia) %in% c(select_var_aic,
                                                                                                       "booking_complete"))], 
                 x = TRUE, y = TRUE, model = TRUE)
lrm.aic 

cross.calib <- calibrate(lrm.aic, method = "crossvalidation", B = 10)
plot(cross.calib, las=1, xlab = "Predicted Probability")
```

AUC and ROC Curve
```{r}
# Predicting probabilities using the logistic regression model
p <- predict(lrm.aic, type = "fitted")

# Generating ROC curve
roc_aic.logit <- roc(customer_booking_malaysia$booking_complete ~ p)

# Extracting True Positive Rate (TPR) and False Positive Rate (FPR)
TPR <- roc_aic.logit$sensitivities
FPR <- 1 - roc_aic.logit$specificities

# Plotting ROC curve
plot(FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2, col = 'red',
     xlab = "False Positive Rate", ylab = "True Positive Rate", main = "ROC Curve")
abline(a = 0, b = 1, lty = 2, col = 'blue')  # Adding diagonal reference line
text(0.7, 0.4, label = paste("AUC =", round(auc(roc_aic.logit), 2)))  # Adding AUC value as text

# Calculating and printing the AUC
auc_value <- auc(roc_aic.logit)
print(paste("AUC value:", round(auc_value, 2)))
```

## 6.1.2 Stepwise AIC Model (Outliers Removed)

Fit logistic regression with the cleaned dataset (without outliers)
```{r, warning=FALSE}
# Combine influential points and outliers without repetition
all_outliers <- unique(outliers_obs)

# Remove outliers and influential points from the dataset
cleaned_data_aic <- customer_booking_malaysia[-all_outliers, ]

# Fit logistic regression model with all variables 
new.logit_model <- glm(booking_complete ~ sales_channel + trip_type + purchase_lead + length_of_stay + flight_hour + wants_preferred_seat + num_passengers + flight_day + route + wants_extra_baggage +
                      wants_in_flight_meals + flight_duration, 
                    family = binomial(link = logit), 
                    data = cleaned_data_aic)
```

Perform Stepwise AIC selection with 'new.logit_model'
```{r}
newsel.var.aic <- step(new.logit_model, trace = 0, k = 2, direction = "both") 
newselect_var_aic <- attr(terms(newsel.var.aic), "term.labels")   
newselect_var_aic
```

Fit Logistic Regression Model with new AIC selection
```{r, warning=FALSE}
new_aic.logit <- glm(booking_complete ~ sales_channel + length_of_stay + flight_duration + num_passengers + wants_extra_baggage + wants_preferred_seat + wants_in_flight_meals, family = binomial(link = logit), data = cleaned_data_aic)
summary(new_aic.logit)
```

Checking influential points
```{r}
# DFBETAS
new_df.aic <- dfbetas(new_aic.logit)
n.aic <- nrow(cleaned_data_aic)
n.beta_cut <- 2 / sqrt(n.aic)
influential_points_naic <- apply(abs(new_df.aic) > n.beta_cut, 1, any)
sum(influential_points_naic)
```

Checking outliers
```{r}
ri.naic <- rstandard(new_aic.logit)
outliers_obs_naic <- which(abs(ri.naic) > 2)
length(outliers_obs_naic)
```

VIF to check for multicollinearity
```{r}
vif(new_aic.logit)
```

Cross-Validation and Calibration
```{r}
set.seed(1007928566)
new_lrm.aic <- lrm(booking_complete ~ ., 
                 data = cleaned_data_aic[, which(colnames(cleaned_data_aic) %in% c(newselect_var_aic,"booking_complete"))], 
                 x = TRUE, y = TRUE, model = TRUE)
new_lrm.aic 

nacross.calib <- calibrate(new_lrm.aic, method = "crossvalidation", B = 10)
plot(nacross.calib, las=1, xlab = "Predicted Probability")
```

AUC and ROC Curve
```{r}
# Predicting probabilities using the logistic regression model
p <- predict(new_lrm.aic, type = "fitted")

# Generating ROC curve
newroc_aic.logit <- roc(cleaned_data_aic$booking_complete ~ p)

# Extracting True Positive Rate (TPR) and False Positive Rate (FPR)
TPR <- newroc_aic.logit$sensitivities
FPR <- 1 - newroc_aic.logit$specificities

# Plotting ROC curve
plot(FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2, col = 'red',
     xlab = "False Positive Rate", ylab = "True Positive Rate", main = "ROC Curve")
abline(a = 0, b = 1, lty = 2, col = 'blue')  # Adding diagonal reference line
text(0.7, 0.4, label = paste("AUC =", round(auc(newroc_aic.logit), 2)))  # Adding AUC value as text

# Calculating and printing the AUC
newauc_value <- auc(newroc_aic.logit)
print(paste("AUC value:", round(newauc_value, 2)))
```

## 6.2.1 Stepwise BIC Model

Fit Logistic Regression Model with BIC selection
```{r, warning=FALSE}
bic.logit <- glm(booking_complete ~ sales_channel + length_of_stay + flight_duration + wants_extra_baggage + wants_preferred_seat + wants_in_flight_meals, family = binomial(link = logit), data = customer_booking_malaysia)
summary(bic.logit)
```

Checking influential points
```{r}
df.bic <- dfbetas(bic.logit)
n.bic <- nrow(customer_booking_malaysia)
beta_cut_bic <- 2 / sqrt(n.bic)
influential_points_bic <- apply(abs(df.bic) > beta_cut_bic, 1, any)
sum(influential_points_bic)
```

Checking outliers
```{r}
ri.bic <- rstandard(bic.logit)
outliers_obs_bic <- which(ri.bic > 2 | ri.bic < -2)
length(outliers_obs_bic )
```

Cross-Validation and Calibration
```{r}
lrm.bic <- lrm(booking_complete ~ ., 
                 data = customer_booking_malaysia[,which(colnames(customer_booking_malaysia) %in% c(select_var_bic, "booking_complete"))], 
                 x = TRUE, y = TRUE, model = TRUE)
lrm.aic 

cross.calib <- calibrate(lrm.bic, method = "crossvalidation", B = 10)
plot(cross.calib, las=1, xlab = "Predicted Probability")
```

VIF to check for multicollinearity
```{r}
vif(bic.logit)
```

AUC and ROC Curve
```{r}
# Predicting probabilities using the logistic regression model
p <- predict(lrm.bic, type = "fitted")

# Generating ROC curve
roc_bic.logit <- roc(customer_booking_malaysia$booking_complete ~ p)

# Extracting True Positive Rate (TPR) and False Positive Rate (FPR)
TPR <- roc_bic.logit$sensitivities
FPR <- 1 - roc_bic.logit$specificities

# Plotting ROC curve
plot(FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2, col = 'red',
     xlab = "False Positive Rate", ylab = "True Positive Rate", main = "ROC Curve")
abline(a = 0, b = 1, lty = 2, col = 'blue')  # Adding diagonal reference line
text(0.7, 0.4, label = paste("AUC =", round(auc(roc_bic.logit), 2)))  # Adding AUC value as text

# Calculating and printing the AUC
bauc_value <- auc(roc_bic.logit)
print(paste("AUC value:", round(bauc_value, 2)))
```

## 6.2.2 Stepwise BIC Model (Outliers removed)

Fit logistic regression with the cleaned dataset (without outliers)
```{r, warning=FALSE}
# Combine influential points and outliers without repetition
all_outliers <- unique(outliers_obs_bic)

# Remove outliers and influential points from the dataset
cleaned_data_bic <- customer_booking_malaysia[-all_outliers, ]

# Fit logistic regression model with all variables 
bnew.logit_model <- glm(booking_complete ~ sales_channel + trip_type + purchase_lead + length_of_stay + flight_hour + wants_preferred_seat + num_passengers + flight_day + route + wants_extra_baggage + wants_in_flight_meals + flight_duration, 
                    family = binomial(link = logit), 
                    data = cleaned_data_bic)
```

Perform Stepwise BIC selection with 'bnew.logit_model'
```{r}
newsel.var.bic <- step(bnew.logit_model, trace = 0, k = log(nrow(cleaned_data_bic)), direction = "both") 
newselect_var_bic <- attr(terms(newsel.var.bic), "term.labels")   
newselect_var_bic
```

Fit Logistic Regression Model with new BIC selection
```{r, warning=FALSE}
new_bic.logit <- glm(booking_complete ~ sales_channel + length_of_stay + flight_duration + wants_extra_baggage + 
                   wants_preferred_seat + wants_in_flight_meals, family = binomial(link = logit),
                 data = cleaned_data_bic)
summary(new_bic.logit)
```

Checking influential points
```{r}
new_df.bic <- dfbetas(new_bic.logit)
nn.bic <- nrow(cleaned_data_bic)
n.beta_cut <- 2 / sqrt(nn.bic)
influential_points_nbic <- apply(abs(new_df.bic) > n.beta_cut, 1, any)
sum(influential_points_nbic)

par(mfrow = c(1, 2))

# Find the index of the predictor variable "length_of_stay"
predictor_index <- which(names(coef(new_bic.logit)) == "length_of_stay")
df.bic_predictor <- new_df.bic[, predictor_index]

# Plot dfbeta against "length_of_stay"
par(family = 'serif')
plot(cleaned_data_bic$length_of_stay, df.bic_predictor, 
     xlab='Length of Stay', ylab='dfbeta', 
     main='Figure 3. DFBETA vs. Length of Stay')
lines(lowess(cleaned_data_bic$length_of_stay, df.bic_predictor), 
      lwd=2, col='blue')
abline(h=0, lty='dotted')
abline(h=-2/sqrt(nrow(new_df.bic)), lty='dotted')
abline(h=2/sqrt(nrow(new_df.bic)), lty='dotted')


# Find the index of the predictor variable "flight_duration"
predictor_index <- which(names(coef(new_bic.logit)) == "flight_duration")
df.bic_predictor <- new_df.bic[, predictor_index]

# Plot dfbeta against "flight_duration"
par(family = 'serif')
plot(cleaned_data_bic$flight_duration, df.bic_predictor, 
     xlab='Flight Duration', ylab='dfbeta', 
     main='DFBETA vs. Flight Duration')
lines(lowess(cleaned_data_bic$flight_duration, df.bic_predictor), 
      lwd=2, col='blue')
abline(h=0, lty='dotted')
abline(h=-2/sqrt(nrow(df.bic)), lty='dotted')
abline(h=2/sqrt(nrow(df.bic)), lty='dotted')

```

Checking outliers
```{r}
ri.nbic <- rstandard(new_bic.logit)
outliers_obs_nbic <- which(ri.nbic > 2 | ri.nbic < -2)
outliers_obs_nbic
```

VIF to check for multicollinearity
```{r}
vif(new_bic.logit)
```

Cross-Validation and Calibration
```{r}
set.seed(1007928566)
new_lrm.bic <- lrm(booking_complete ~ ., 
                 data = cleaned_data_bic[, which(colnames(cleaned_data_bic) %in% c(newselect_var_bic, "booking_complete"))], 
                 x = TRUE, y = TRUE, model = TRUE)
new_lrm.bic 

nbcross.calib <- calibrate(new_lrm.bic, method = "crossvalidation", B = 10)
plot(nbcross.calib, las=1, xlab = "Predicted Probability", main = "Figure 1. The Calibration Plot")

```

AUC and ROC Curve
```{r}
# Predicting probabilities using the logistic regression model
p <- predict(new_lrm.bic, type = "fitted")

# Generating ROC curve
newroc_bic.logit <- roc(cleaned_data_bic$booking_complete ~ p)

# Extracting True Positive Rate (TPR) and False Positive Rate (FPR)
TPR <- newroc_bic.logit$sensitivities
FPR <- 1 - newroc_bic.logit$specificities

# Plotting ROC curve
plot(FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2, col = 'red',
     xlab = "False Positive Rate", ylab = "True Positive Rate", main = "Figue 2. ROC Curve")
abline(a = 0, b = 1, lty = 2, col = 'blue')  # Adding diagonal reference line
text(0.7, 0.4, label = paste("AUC =", round(auc(newroc_bic.logit), 2)))  # Adding AUC value as text

# Calculating and printing the AUC
nbauc_value <- auc(newroc_bic.logit)
print(paste("AUC value:", round(nbauc_value, 2)))
```

# Exploratory Data Analysis of Stepwise BIC with cleaned dataset (Chosen Model)
```{r}
# Set up 1x2 plotting window
par(mfrow = c(2, 4))

# Histograms for numerical variables
hist(cleaned_data_bic$flight_duration,
     main = "Hist. Flight Duration",
     xlab = "Flight Duration",
     col = "orange")

hist(cleaned_data_bic$length_of_stay,
     main = "Hist. of Length of Stay",
     xlab = "Length of Stay",
     col = "orange")

# Bar plots for categorical variables
barplot(table(cleaned_data_bic$sales_channel),
        main = "Sales Channel Dist.",
        col = c("lavender", "lightblue")) 

barplot(table(cleaned_data_bic$wants_extra_baggage),
        main = "Extra Baggage Dist.",
        col = c("lavender", "lightblue"))

barplot(table(cleaned_data_bic$wants_preferred_seat),
        main = "Preferred Seat Dist.",
        col = c("lavender", "lightblue"))

barplot(table(cleaned_data_bic$wants_in_flight_meals),
        main = "In-Flight Meals Dist.",
        col = c("lavender", "lightblue"))

barplot(table(cleaned_data_bic$booking_complete),
        main = "Booking Complete Dist.",
        col = c("lavender", "lightblue"))
```

```{r}
# Create the contingency table for categorical variable (sales_channel)
bivariate_table <- table(cleaned_data_bic$sales_channel, cleaned_data_bic$booking_complete)

# Add margins (totals) to the table
bivariate_table_with_margins <- addmargins(bivariate_table)

# Rename the last column to "Total"
colnames(bivariate_table_with_margins)[ncol(bivariate_table_with_margins)] <- "Total"

# Rename the last row to "Total"
rownames(bivariate_table_with_margins)[nrow(bivariate_table_with_margins)] <- "Total"

# Print the table with renamed margins
print(bivariate_table_with_margins)
```

```{r}
# Create the contingency table for categorical variables (wants_extra_baggage)
bivariate_table <- table(cleaned_data_bic$wants_extra_baggage, cleaned_data_bic$booking_complete)

# Add margins (totals) to the table
bivariate_table_with_margins <- addmargins(bivariate_table)

# Rename the last column to "Total"
colnames(bivariate_table_with_margins)[ncol(bivariate_table_with_margins)] <- "Total"

# Rename the last row to "Total"
rownames(bivariate_table_with_margins)[nrow(bivariate_table_with_margins)] <- "Total"

# Print the table with renamed margins
print(bivariate_table_with_margins)
```

```{r}
# Create the contingency table for categorical variables (wants_preferred_seat)
bivariate_table <- table(cleaned_data_bic$wants_preferred_seat, cleaned_data_bic$booking_complete)

# Add margins (totals) to the table
bivariate_table_with_margins <- addmargins(bivariate_table)

# Rename the last column to "Total"
colnames(bivariate_table_with_margins)[ncol(bivariate_table_with_margins)] <- "Total"

# Rename the last row to "Total"
rownames(bivariate_table_with_margins)[nrow(bivariate_table_with_margins)] <- "Total"

# Print the table with renamed margins
print(bivariate_table_with_margins)
```

```{r}
# Create the contingency table for categorical variables (wants_in_flight_meals)
bivariate_table <- table(cleaned_data_bic$wants_in_flight_meals, cleaned_data_bic$booking_complete)

# Add margins (totals) to the table
bivariate_table_with_margins <- addmargins(bivariate_table)

# Rename the last column to "Total"
colnames(bivariate_table_with_margins)[ncol(bivariate_table_with_margins)] <- "Total"

# Rename the last row to "Total"
rownames(bivariate_table_with_margins)[nrow(bivariate_table_with_margins)] <- "Total"

# Print the table with renamed margins
print(bivariate_table_with_margins)
```


